Expert-In-The-Loop Causal Discovery: \\ Iterative Model Refinement Using Expert Knowledge

### 2025-02-10 paper/main.tex
Updates background

\subsection{Testing DAGs Using CI Tests}
Our structure learning algorithm uses the idea of testing DAGs using CI tests.
In this section, we give an example of how this model testing works in
practice. We take the example of the MEC learned by the PC algorithm in
Figure~\ref{fig:intro}(a) using $ N=800 $. We first orient all the undirected
edges in the model based on our domain knowledge to obtain a DAG. In this case
we only have one undirected edge between Marital Status and Relationship. We 
orient the edge as Marital Status to Relationship. We then proceed to test
each of the implied CI of the model using a statistical test. We show the result
of some of these tests in Figure.

@@

Polyserial (when one is ordinal and the other is continuous) and Polychoric
(when both are ordinal) correlation have been used to estimate the covariance
matrix between them \citep{Poon1987}. Both methods make the assumption that the
observed ordinal variable is a result of thresholding a latent normally
distributed continuous variable. Under this assumption, the methods then try to
estimate the covariance matrix while maximizing the likelihood of the dataset.
Using the estimated covariance matrix, $ \Sigma $ we can the compute Pearson's
correlation coefficient to use as the effect size as well as for testing. When
$ \bm{Z} = \emptyset $,

\begin{equation}
	r_{X, Y} = \frac{\Sigma_{XY}}{\sqrt{\Sigma_{XX} \Sigma_{YY}}}
\end{equation}

When $ \bm{Z} \neq \emptyset $,

\begin{equation}
	r_{X, Y \rvert \bm{Z}} = - \frac{\Sigma^{-1}_{XY}}{\sqrt{\Sigma^{-1}_{XX} \Sigma^{-1}_{YY}}}
\end{equation}

@@

In the case when we have a combination of ordinal and categorical variables, we
can use the residualization based CI test from \citep{Ankan2023}. This test
gives a chi-square distributed test statistic, that can be used for computing
an effect size such as Root Mean Squared Error of Approximation (RMSEA). Given
a chi-squared test statistic, $ \chi^2 $, with $ \textrm{df} $ degrees of
freedom, the RMSEA is defined as:

>>>>

\subsection{Testing DAG Using CI Tests}
Our structure learning algorithm uses the idea of testing DAGs using CI tests.
In this section, we give an example of how this model testing works in
practice. We take the example of the MEC learned by the PC algorithm in
Figure~\ref{fig:intro}(a) using $ N=800 $. We first orient all the undirected
edges in the model based on our domain knowledge to obtain a DAG. In this case
we only have one undirected edge between Marital Status and Relationship. We 
orient the edge as Marital Status to Relationship. We then proceed to test
each of the implied CI of the model using a statistical test. We show the result
of some of these tests in Figure.

\todo[inline]{Add a table here}
% \begin{figure}
% 	\centering
% 	\begin{tabular}{llrr}
% 		\multicolumn{4}{c}{Z=Age, Sex} \\
% 		$X$ & $Y$ & \bm{Z} & p \\
% 		\hline
% 		Edct & Wrkc & 1.00  \\
% 		Occp & Wrkc & 1.00  \\
% 		Rltn & HrPW &  .99  \\
% 		Incm & Occp & .08   \\
% 		Incm & Wrkc &  .50  \\
% 		Incm & HrPW & .003  \\
% 		\hline
% 	\end{tabular}
% \end{figure}

@@

Polyserial (when one is ordinal and the other is continuous) and Polychoric
(when both are ordinal) correlation have been used to estimate the covariance
matrix between them \citep{Poon1987}. Both methods make the assumption that the
observed ordinal variable is a result of thresholding a latent normally
distributed continuous variable. Under this assumption, the methods then try to
estimate the covariance matrix while maximizing the likelihood of the dataset.
Using the estimated covariance matrix, $ \Sigma $ we can perform the CI test and compute 
the Pearson's correlation coefficient as the effect size:

\begin{equation}
	r_{X, Y \rvert \bm{Z} = \emptyset} = \frac{\Sigma_{XY}}{\sqrt{\Sigma_{XX} \Sigma_{YY}}}; \;\;\; \; r_{X, Y \rvert \bm{Z}} = - \frac{\Sigma^{-1}_{XY}}{\sqrt{\Sigma^{-1}_{XX} \Sigma^{-1}_{YY}}}
\end{equation}

@@

In the case when we have a combination of ordinal and categorical variables, we
can use a residualization based CI test \citep{Ankan2023}. This test gives a
chi-square distributed test statistic, that can be used for computing an effect
size such as the Root Mean Squared Error of Approximation (RMSEA). Given a
chi-squared test statistic, $ \chi^2 $, with $ \textrm{df} $ degrees of
freedom, the RMSEA is defined as:
